
\documentclass{beamer}

\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{amssymb}
\begin{document}


\begin{frame}
	\frametitle{Example 2: Evaluating a Function}
	Evaluate the following function for x =1,2 and 5 respectively.
	{
		\Large
	\[ f(x) = \frac{e^x + {e^{-x}}}{2} \]
}
	Remark:  Example 1 was done in previous class.
\end{frame}
%================================================================================== %
\begin{frame}
	
	\frametitle{Example 3: Evaluating a Function}
	Evaluate the function for each of the following values : 0.5,1,1.25,2.
	{
		\Large
	\[f(x) =  \sqrt{1+e^{x}}  \]
}
	Four decimal places will suffice.
\end{frame}
\begin{frame}
	\begin{center}
		\begin{tabular}{|c|c|c|c|}
			\hline \rule[-2ex]{0pt}{5.5ex} x & $e^x$ & $1+e^x$ & $\sqrt{1+e^x}$ \\ \hline
			\hline \rule[-2ex]{0pt}{5.5ex} 0.5 & \phantom{space}  &  &  \\ 
			\hline \rule[-2ex]{0pt}{5.5ex} 1 &  &  &  \\ 
			\hline \rule[-2ex]{0pt}{5.5ex} 1.25 & \phantom{space}  &  &  \\ 
			\hline \rule[-2ex]{0pt}{5.5ex} 2 & \phantom{space}  &  & \phantom{space}  \\ 
			\hline 
		\end{tabular} 
	\end{center}
\end{frame}
%================================================================================ %
\begin{frame}
	\frametitle{Recall : Sets of Numbers}
	\begin{itemize} 
		\item $\mathbb{N}$ Set of all natural numbers
		\item $\mathbb{Z}$ Set of all integers
		\item $\mathbb{Q}$ Set of all rational numbers
		\item $\mathbb{R}$ Set of all real numbers
	\end{itemize}
	
	\bigskip There are, of course, other numbers sets, but we will not be encountering them on the course.
\end{frame}
%========================================================== %

\begin{frame}
	\frametitle{Recall : Sets of Numbers}
	\begin{itemize}
		\item $\mathbb{Z}^{+}$ Set of all positive integers
		\item $\mathbb{Z}^{-}$ Set of all negative integers
		\item $\mathbb{R}^{+}$ Set of all positive real numbers
		\item $\mathbb{R}^{-}$ Set of all negative real numbers
	\end{itemize}
\end{frame}

%=========================================================================================== %
\subsection*{Topic 1 : Special Functions}
\begin{frame}
	\frametitle{Special Functions}
	\begin{itemize}
		% \item The factorial Operator
		\item Absolute Value Function
		\item The Sign Function
		\item Floor and Ceiling Functions
		\item Hyperbolic Functions
	\end{itemize}	
\end{frame}


%---------------------------------------%
\begin{frame}
	\frametitle{Absolute Value Function}
	\Large
	\textbf{Absolute Value Function}
	\begin{itemize}
		\item The absolute value (or modulus) $|x|$ of a real number x is the non-negative value of x without regard to its sign. 
		
		%	\item For example, $|x| = x$ for a positive x, $|x| = -x$ for a negative x (in which case −x is positive), and $|0| = 0$. 
		%	\item For example, the absolute value of 4 is 4, and the absolute value of $-4$ is also 4. 
		%\item The absolute value of a number may be thought of as its distance from zero.
	\end{itemize}
	\[|x| = \begin{cases} x, & \mbox{if }  x \ge 0  \\ -x,  & \mbox{if } x < 0. \end{cases} \]
\end{frame}
%---------------------------------------%
\begin{frame}
	\frametitle{Topic 1 : Absolute Value Function}
	\Large
	\vspace{-1cm}
	\begin{itemize}
		
		\item For a positive x, $|x| = x$ 
		\item For a negative x (in which case −x is positive) $|x| = -x$ 
		\item The absolute value of 0 is 0:  $|0| = 0$. 
		\item For example, the absolute value of 4 is 4, and the absolute value of $-4$ is also 4. 
		\item IMPORTANT:  The input to this function is any real number. The output of this function will always be a positive real numbers.
		%\item The absolute value of a number may be thought of as its distance from zero.
	\end{itemize}
	
	
\end{frame}
\begin{frame}
	\frametitle{Topic 1 : Sign Function}
	\Large
	\textbf{Sign Function}
	\begin{itemize}
		\item The sign function $sng(x)$ of a real number x is a signed value of absolute value of 1, dependent on the sign of $x$. 
		
		%	\item For example, $|x| = x$ for a positive x, $|x| = -x$ for a negative x (in which case −x is positive), and $|0| = 0$. 
		%	\item For example, the absolute value of 4 is 4, and the absolute value of $-4$ is also 4. 
		%\item The absolute value of a number may be thought of as its distance from zero.
	\item IMPORTANT:  The input to this function is any real number. The output of this function will always be either 1 or -1.
	\end{itemize}
	\[ sgn(x) = \begin{cases} 1, & \mbox{if }  x \ge 0  \\ -1,  & \mbox{if } x < 0. \end{cases} \]
\end{frame}
%---------------------------------------%
\begin{frame}
	\frametitle{Topic 1 : Floor and Ceiling Functions}
	\Large
	\begin{itemize}
		\item The floor and ceiling functions map a real number to the largest previous or the smallest following integer, respectively. \item More precisely, \[floor(x) = \lfloor x\rfloor \] is the largest integer not greater than x and \[ceiling(x) =  \lceil x \rceil \] is the smallest integer not less than x.
	\end{itemize}
	
	
\end{frame}

%---------------------------------------%
\begin{frame}
	\frametitle{Topic 1 : Floor and Ceiling Functions}
	\textbf{Examples}
	{
		\LARGE
		\begin{eqnarray}
		\lfloor 3.14 \rfloor =& 3 \\ \bigskip
		\lceil -4.5 \rceil =& -5 \\ \bigskip
		| -4 | =&  4
		\end{eqnarray}
}
{\Large	
	\textbf{Remark:}  Input to the floor and ceiling function can be any really number, but outputs are always integers.
	}
\end{frame}


\begin{frame}
	\frametitle{ Topic 2 : Laws of Logarithms}
	\Large
	\begin{itemize}
		\item Law 1 : Multiplcation of Logarithms
		\[ Log(a) \times Log(b) = Log(a+b) \]
		\item Law 2 : Division of Logarithms
		\[ \frac{Log(a)}{Log(b)} = Log(a-b) \]
		\item Law 3 : Powers of Logarithms
		\[ Log(a^b) = b \times Log(a) \]
	\end{itemize}
	
	
	
\end{frame}
%======================================================================================================== %

\begin{frame}
	\frametitle{ Topic 2 : Laws of Logarithms}
	\Large
	\begin{itemize}
		\item Law 1 : Multiplcation of Logarithms
		\[ Log(a) \times Log(b) = Log(a+b) \]
		\item Law 2 : Division of Logarithms
		\[ \frac{Log(a)}{Log(b)} = Log(a-b) \]
		\item Law 3 : Powers of Logarithms
		\[ Log(a^b) = b \times Log(a) \]
	\end{itemize}
	
	
	
\end{frame}
%========================================================================================================== %
\end{document}

