\documentclass[a4paper,12pt]{article}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage{eurosym}
\usepackage{vmargin}
\usepackage{amsmath}
\usepackage{graphics}
\usepackage{epsfig}
\usepackage{subfigure}
\usepackage{fancyhdr}

\setcounter{MaxMatrixCols}{10}
%TCIDATA{OutputFilter=LATEX.DLL}
%TCIDATA{Version=5.00.0.2570}
%TCIDATA{<META NAME="SaveForMode" CONTENT="1">}
%TCIDATA{LastRevised=Wednesday, February 23, 2011 13:24:34}
%TCIDATA{<META NAME="GraphicsSave" CONTENT="32">}
%TCIDATA{Language=American English}

\pagestyle{fancy}
\setmarginsrb{20mm}{0mm}{20mm}{25mm}{12mm}{11mm}{0mm}{11mm}
\lhead{Hibernia College} \rhead{Kevin O'Brien}
\chead{Mathematics for Computing Part 2 Sheet A}
%\input{tcilatex}

\begin{document}

\section*{Rules of Inclusion, Listing and Cardinality}
For each of the following sets, a set is specified by the rules of inclusion method and listing method respectively. Also stated is the cardinality of that data set.
\subsection*{Worked example 1}
\begin{itemize}
\item $\{ x : x $ is an odd integer $ 5 \leq x \leq 17 \}$
\item $x = \{5,7,9,11,13,15,17\}$
\item The cardinality of set $x$ is 7.
\end{itemize}

\subsection*{Worked example 2}
\begin{itemize}
\item $\{ y : y $ is an even integer $ 6 \leq y < 18 \}$
\item $y = \{6,8,10,12,14,16\}$
\item The cardinality of set $y$ is 6.
\end{itemize}

\subsection*{Worked example 3}
A perfect square is a number that has a integer value as a
square root. 4 and 9 are perfect squares ($\sqrt{4} = 2$,
$\sqrt{9} = 3$).
\begin{itemize}
\item $\{ z : z $ is an perfect square $ 1 < z < 100 \}$
\item $z = \{4,9,16,25,36,49,64,81\}$
\item The cardinality of set $z$ is 8.
\end{itemize}


\subsection*{Exercises}
For each of the following sets, write out the set using the listing method.
Also write down the cardinality of each set.

\begin{itemize}
\item $\{ s : s $ is an negative integer $ -10 \leq s \leq 0 \}$
\item $\{ t : t $ is an even number $ 1 \leq t \leq 20 \}$
\item $\{ u : u $ is a prime number $ 1 \leq u \leq 20 \}$
\item $\{ v : v $ is a multiple of 3 $ 1 \leq v \leq 20 \}$
\end{itemize}
%-------------------------------------------------%

\section*{Power Sets}
\subsection*{Worked Example}
Consider the set $Z$:
\[ Z = \{ a,b,c\}  \]
\begin{itemize}
\item[Q1] How many sets are in the power set of $Z$?
\item[Q2] Write out the power set of $Z$.
\item[Q3] How many elements are in each element set?
\end{itemize}
%----------------------------------------------%
\subsection*{Solutions to Worked Example}

\begin{itemize}


\item[Q1] There are 3 elements in $Z$. So there is $2^3 = 8$ element sets contained in the power set.

\item[Q2] Write out the power set of $Z$.
\[ \mathcal{P}(Z) = \{ \{0\}, \{a\}, \{b\}, \{c\}, \{a,b\}, \{a,c\}, \{b,c\}, \{a,b,c\} \]

\item[Q3]
\begin{itemize}
\item[*] One element set is the null set - i.e. containing no
elements \item[*] Three element sets have only elements \item[*]
Three element sets have two elements \item[*] One element set
contains all three elements \item[*] 1+3+3+1=8
\end{itemize}
\end{itemize}
\subsection*{Exercise}
For the set $Y = \{u,v,w,x\}$ , answer the questions from the
previous exercise


%------------------------------------------------------%

\section*{Complement of a Set}
%(2.3.1)
Consider the universal set $U$ such that
\[U=\{2,4,6,8,10,12,15\} \]
For each of the sets $A$,$B$,$C$ and $D$, specify the complement sets.

\begin{center}
\begin{tabular}{|c|c|}
  \hline
Set & Complement\\
\hline $A=\{4,6,12,15\}$ &
$A^{\prime}=\{2,8,10\}$ \\ \hline $B=\{4,8,10,15\}$ & \\ \hline
$C=\{2,6,12,15\}$ & \\ \hline $D=\{8,10,15\}$ & \\ \hline

\end{tabular}
\end{center}
%-------------------------------------%
 % Binary Operations on Sets (2.3.2)
 % Union , Intersection, Symmetric Difference
 % Set Difference



\section*{Set Operations}
Consider the universal set $U$ such that
\[U=\{1,2,3,4,5,6,7,8,9\} \]
and the sets
\[A=\{2,5,7,9\} \]
\[B=\{2,4,6,8,9\} \]

\begin{itemize}
\item[(a)] $A-B$
\item[(b)] $A \otimes B$
\item[(c)] $A \cap B$
\item[(d)] $A \cup B$
\item[(e)] $A^{\prime} \cap B^{\prime}$
\item[(f)] $A^{\prime} \cup B^{\prime}$
\end{itemize}


\section*{Venn Diagrams}

Draw a Venn Diagram to represent the universal set
$\mathcal{U} = \{0,1,2,3,4,5,6\}$ with subsets
$A = \{2,4,5\}$
$B = \{1,4,5,6\}$

Find each of the following
\begin{itemize}
\item[(a)] $A \cup B $
\item[(b)] $A \cap B $
\item[(c)] $A-B$
\item[(d)] $B-A$
\item[(e)] $A \otimes B$
\end{itemize}

\newpage
\subsection*{Conversion to Binary Form}
\[
\begin{tabular}{|c|c|c|c|}
	&		&	Quotient	&	Remainder	\\
507	&\emph{	253.5	}&	253	&	1	\\
253	&\emph{	126.5	}&	126	&	1	\\
126	&\emph{	63	}&	63	&	0	\\
63	&\emph{	31.5	}&	31	&	1	\\
31	&\emph{	15.5	}&	15	&	1	\\
15	&\emph{	7.5	}&	7	&	1	\\
7	&\emph{	3.5	}&	3	&	1	\\
3	&\emph{	1.5	}&	1	&	1	\\
1	&\emph{	0.5	}&	0	&	1	\\
\end{tabular} \]

Correct Answer : 111111011


2011 Question 1C

List all the set of positive integers with precisely 3 bits in binary notation

\[\begin{array}{|c|c|}
100 & 4 \\
101 & 5 \\
110 & 6 \\
111 & 7 \\
\end{array}\]

Let n be a positive integer. How many positive integers have precisely n bits in binary notation. 

\newpage
Express the binary number (1011.011) as a decimal, showing all of your working
\[\begin{tabular}{|c|c|c|c|}
Digit	&	Power	&	Weight	&		\\
1	&	3	&	2^{3} = 8	&	8	\\
0	&	2	&	2^{2} =4	&	0	\\
1	&	1	&	2^{1}=2	&	2	\\
1	&	0	&	2^{0}=1	&	1	\\
.	&	.	&	.	&	.	\\
0	&	-1	&	2^{-1}=0.5	&	0	\\
1	&	-2	&	2^{-1}=0.25	&	0.25	\\
1	&	-3	&	2^{-3}=0.125	&	0.125	\\
	&		&		&	11.375	\\
\hline 
\end{tabular} \]





\newpage
\subsection*{2011 Question 1b}

Working in Binary, and showing all carries, compute the following:\\
$(11010)_{2} + (111)_{2}$ (i.e. $(26)_10 +7_{10}$)
\\
(Demonstration on whiteboard)
\\
Answer: 11010 + 111 = 100001
\\


\begin{itemize} 
\item Remark 0 is not a positive integer.
\item How many have 1 bit : Answer only one 1.
\item How many have 2 bits :  Answer only two : 10 and 11 (i.e. 2 and 3 in decimal)
\item From above: 4 have 3 bits.
\item What is the first 4 bit numbers? : 1000 (i.e 8 in decimal)
\item What is the last 4 bit number? 1111 (i.e 15 in decimal)
\item How many numbers have 4 bits? Answer :8
\item Answer to question: $2^{n-1}$
\item Return to this after "Proof by Induction".
\end{itemize}
\newpage

\newpage
\begin{verbatim}
If U = {2, 3, 4, 7, 9, 10, 11, 13} and A = {3, 4, 9, 10}

Then

Compliment of A = A’	 = U - A
= {2, 3, 4, 7, 9, 10, 11, 13} – {3, 4, 9, 10}
= {2, 7, 11, 13}

\begin{verbatim}
UNIVERSAL SET: 

A universal set is a set of all elements under consideration. It is denoted by U. 
A Universal set is always a non-empty set.

For example: The set of real numbers R is a universal set
for the operations related to real numbers.

Example: 

Given that U = {5, 6, 7, 8, 9, 10, 11, 12}, list the elements of the following sets. 
A = {x : x is a prime number}
B = {x : x is a factor of 60}
Solution: 

The elements of sets A and B can only be selected 
from the given universal set U 
A = {5, 7, 11}
B = {5, 6, 10, 12}
\end{verbatim}

\newpage
\begin{verbatim}
MUTUALLY EXCLUSIVE SETS OR DISJOINT SETS: 

Two sets are called disjoint if they don't have any common element. 

For example, A = {2, 3, 4} and B = {5, 6, 7} are disjoint sets.

\end{verbatim}

\end{document}








%-------------------------------------% Ellipsis

When using Ellipsis, it should be clear what the pattern is

%-------------------------------------%


%-----------Reference to section 2.2.3 Power Sets




%-------------------------------------%

 %----(Reference to Section 2.2.2 Cardinality)



%-------------------------------------% % Complement of a set


%--------------------------------------%
\subsection*{ Three Sets }

%- Section 2.3.5 %- Associative Law %- Distribution Law





%-------------------------------------% %- Section 3
Propositional Logic A statement is a declarative sentence that
is either true or false.
\begin{itemize}
\item $\tilde q$ not q \item $p \vee q$ \item $p \wedge \tilde
q$
\end{itemize}

%-------------------------------%
------------------------------------------------
Question 1


$\{ s :  s is an odd integer and 2 \leq s \leq 10 \}$
$\{ 2m :  m \element Z and 5 \leq m \leq 10 \}$
$\{ 2^t :  t \element Z and 0 \leq t \leq 5 \}$

------------------------------------------------
Question 2

\{12,13,14,15,16,17\}
{0,5,-5,10,-10,15,-15,.....\}
{6,8,10,12,14,16,18\}
Question 5

-----------------------------------------------------


--------------------------------------------

Let A, B be subsets of the universal set \mathcal{U}.

Use membership tables to prove De Morgan's Laws.

---------------------------------------------
Draw a venn diagram to show three subsets A,B and C of a universal set U intersecting in
the most general way?
How are sets $X$ and $Z$ related?
Can you describe each of the subsets X,Y and Z in terms  of the
sets A,B,C using the operations union intersection and set complement.

Construct Membership tables for each of the sets
(A-B) - C
A-(B- C)

(A-B) -C = A-(B-C)
A



\begin{itemize}
\item[a.] (1 mark) Write out the sample space for the outcomes for both players A and B.
\item[b.] (1 mark) Write out the sample space for the outcomes of C, where C is the difference of the two scores (i.e. B-A)
\item[c.] (1 mark) Are the sample points for the sample space of C equally probable? Provide a brief justification for your answer.
\end{itemize}

%----------------------------------------------------------%
\newpage
\section*{Section B: Set Operations}
\begin{itemize}
\item[B.1] complement of A $A^{\prime}$
\item[B.2] Union $A \cup B$
\item[B.3] Intersection $A \cap B$
\item[B.4] Relative Difference $A \otimes B$
\item[A.5]
\item[A.6]
\item[A.7]
\item[A.8]
\end{itemize}
\newpage


\begin{itemize}
\item Specifying Sets
\item Listing Method
\item Rules of Inclusion method
\end{itemize}


\begin{itemize}
\item Subsets Notation of a subset
\item Cardinality of a set
\item Power of a set
\end{itemize}

\subsection*{Operation on Sets}

\begin{itemize}
\item The complement of Set
\item Binary Operations
\begin{itemize}
\item Union
\item Intersection
\end{itemize}
\item Membership tables
\item Laws for Combining Sets
\end{itemize}
%----------------------------------------------------------%
\subsection*{Membership Tables}
Using membership tables
\begin{tabular}{|ccc|c|c|c|}
  \hline
  % after \\: \hline or \cline{col1-col2} \cline{col3-col4} ...
  A & B & C & x & y & z \\\hline
  0 & 0 & 0 & 1 & 1 & 1 \\
  0 & 0 & 1 & 0 & 0 & 1 \\
  0 & 1 & 0 & 0 & 0 & 1 \\
  0 & 1 & 1 & 0 & 0 & 1 \\
  1 & 0 & 0 & 1 & 0 & 1 \\
  1 & 0 & 1 & 1 & 0 & 1 \\
  1 & 1 & 0 & 0 & 0 & 1 \\
  1 & 1 & 1 & 1 & 0 & 1 \\
  \hline
\end{tabular}

%